Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond

### 2024-02-13 assets/simulation/pc_singular_extreme.tex
:paintbrush: Update plot functions

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44700,0.74100}%
\definecolor{mycolor2}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor3}{rgb}{0.85098,0.32549,0.09804}%
\definecolor{mycolor4}{rgb}{0.49400,0.18400,0.55600}%
\definecolor{mycolor5}{rgb}{0.46667,0.67451,0.18824}%
%
\makeatletter
\newcommand\resetstackedplots{
    \makeatletter
    \pgfplots@stacked@isfirstplottrue
    \makeatother
    \addplot [forget plot,draw=none] coordinates{(1,0) (5,0) (10,0) (15,0) (20,0)};
}
\makeatother
%
\begin{tikzpicture}

\begin{axis}[%
width=9.509cm,
height=7.5cm,
at={(0cm,0cm)},
scale only axis,
bar width=0.325,
xmin=-0.125,
xmax=5.125,
xtick={1,2,3,4},
xticklabels={{$\sigma_1(\mathbf{H})$},{$\sigma_2(\mathbf{H})$},{$\sigma_3(\mathbf{H})$},{$\sigma_4(\mathbf{H})$}},
ymin=0,
ymax=0.003,
ylabel style={font=\color{white!15!black}},
ylabel={Amplitude},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
legend style={legend cell align=left, align=left, draw=white!15!black},
legend columns=4,
transpose legend,
legend style={/tikz/column 2/.style={column sep=5pt}}
]
\addplot[ybar stacked, fill=mycolor1, fill opacity=0, draw=none, forget plot] table[row sep=crcr] {%
0.8375	0.00195358578718362\\
1.8375	0.000905403110314898\\
2.8375	0.000478480067807894\\
3.8375	3.24792561983075e-17\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addplot[ybar stacked, fill=mycolor2, fill opacity=0.5, draw=black, area legend] table[row sep=crcr] {%
0.8375	4.10673634284198e-05\\
1.8375	9.08396603062443e-05\\
2.8375	0.000280905224832324\\
3.8375	0.000316345482660046\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addlegendentry{D-min}

\addplot[ybar stacked, fill=mycolor3, fill opacity=0.5, draw=black, area legend] table[row sep=crcr] {%
0.8375	0.00057247498440489\\
1.8375	0.000934081190202522\\
2.8375	0.00014601781767251\\
3.8375	0.000162134585145759\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addlegendentry{D-max}

\resetstackedplots

\addplot[ybar stacked, fill=mycolor4, fill opacity=0, draw=none, forget plot] table[row sep=crcr] {%
1.1625	0.00195358621182515\\
2.1625	0.000905405810406567\\
3.1625	0.000478484657925968\\
4.1625	3.62951302572638e-06\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addplot[ybar stacked, fill=mycolor2, draw=black, area legend] table[row sep=crcr] {%
1.1625	4.10669387868874e-05\\
2.1625	9.0836960214575e-05\\
3.1625	0.000280900634714249\\
4.1625	0.000312715969634352\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addlegendentry{BD-min}

\addplot[ybar stacked, fill=mycolor3, draw=black, area legend] table[row sep=crcr] {%
1.1625	0.000668204539596567\\
2.1625	0.000957343016500992\\
3.1625	0.000146017817609404\\
4.1625	0.000162134585134279\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addlegendentry{BD-max}

\addplot [color=mycolor5, line width=2.0pt]
  table[row sep=crcr]{%
-0.125	0.00195358578716264\\
5.125	0.00195358578716264\\
};
\addlegendentry{$\sigma_1(\mathbf{T})$}

\addplot [color=mycolor5, dashed, line width=2.0pt]
  table[row sep=crcr]{%
-0.125	0.000905403110313456\\
5.125	0.000905403110313456\\
};
\addlegendentry{$\sigma_2(\mathbf{T})$}

\addplot [color=mycolor5, dotted, line width=2.0pt]
  table[row sep=crcr]{%
-0.125	0.000478480067806991\\
5.125	0.000478480067806991\\
};
\addlegendentry{$\sigma_3(\mathbf{T})$}

\addplot [color=mycolor5, dashdotted, line width=2.0pt]
  table[row sep=crcr]{%
-0.125	1.20606340074858e-11\\
5.125	1.20606340074858e-11\\
};
\addlegendentry{$\sigma_4(\mathbf{T})$}

>>>>

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44700,0.74100}%
\definecolor{mycolor2}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor3}{rgb}{0.85098,0.32549,0.09804}%
\definecolor{mycolor4}{rgb}{0.49400,0.18400,0.55600}%
%
\begin{tikzpicture}

\begin{axis}[%
width=9.509cm,
height=7.5cm,
at={(0cm,0cm)},
scale only axis,
bar width=0.325,
xmin=-0.125,
xmax=5.125,
xtick={1,2,3,4},
xticklabels={{$\sigma_1(\mathbf{H})$},{$\sigma_2(\mathbf{H})$},{$\sigma_3(\mathbf{H})$},{$\sigma_4(\mathbf{H})$}},
ymin=0,
ymax=0.003,
ylabel style={font=\color{white!15!black}},
ylabel={Amplitude},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
legend style={legend cell align=left, align=left, draw=white!15!black},
legend columns=4,
transpose legend,
legend style={/tikz/column 2/.style={column sep=5pt}}
]
\addplot[ybar stacked, fill=mycolor1, fill opacity=0, draw=black, area legend] table[row sep=crcr] {%
0.8375	0.00143117804054061\\
1.8375	0.000743199099777249\\
2.8375	0.000269283237979035\\
3.8375	8.24218765062706e-09\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addplot[ybar stacked, fill=mycolor2, fill opacity=0.5, draw=black, area legend] table[row sep=crcr] {%
0.8375	0.000462878936482017\\
1.8375	0.00042549634196207\\
2.8375	0.00041106144944578\\
3.8375	0.00024003624027829\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addlegendentry{D-min}

\addplot[ybar stacked, fill=mycolor3, fill opacity=0.5, draw=black, area legend] table[row sep=crcr] {%
0.8375	0.000506636656075082\\
1.8375	0.000495702837065026\\
2.8375	0.000451079141362691\\
3.8375	0.000460530423624515\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addlegendentry{D-max}

\addplot[ybar stacked, fill=mycolor4, fill opacity=0, draw=black, area legend] table[row sep=crcr] {%
1.1625	0.00129022242555956\\
2.1625	0.000545548715650276\\
3.1625	8.37754891714641e-05\\
4.1625	1.02575159048296e-08\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addplot[ybar stacked, fill=mycolor2, draw=black, area legend] table[row sep=crcr] {%
1.1625	0.000603834551463065\\
2.1625	0.000623146726089043\\
3.1625	0.000596569198253351\\
4.1625	0.000240034224950036\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addlegendentry{BD-min}

\addplot[ybar stacked, fill=mycolor3, draw=black, area legend] table[row sep=crcr] {%
1.1625	0.000631314400677068\\
2.1625	0.00064060657290608\\
3.1625	0.000623100053838356\\
4.1625	0.000631935100972121\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
-0.125	0\\
5.125	0\\
};
\addlegendentry{BD-max}

### 2024-02-13 main.tex
:paintbrush: Update plot functions

\begin{section}{Channel Singular Values Redistribution}
	\begin{subsection}{A Toy Example}
		We first illustrate the channel shaping capabilities of different \gls{ris} by a toy example.
		% We first use a toy example to illustrate that \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
		Consider a $2 \times 2 \times 2$ setup where the direct link is blocked.
		The diagonal \gls{ris} is modeled by $\mathbf{\Theta}_\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{bd}-\gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}_\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		In particular, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		We also enforce symmetry by $\beta = \pi / 2$ such that both architectures have the same number of angular parameters.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/simulation/pc_singular_toy.eps}
			\caption{$2 \times 2 \times 2$ channel singular value shaping by diagonal and symmetry unitary \gls{ris}. Direct link is absent.}
			\label{sm:pc_singular_toy}
		\end{figure}
		Fig.~\ref{sm:pc_singular_toy} shows the channel singular values achieved by an exhaustive grid search over $(\theta_1, \theta_2)$ for diagonal \gls{ris} and $(\alpha, \psi)$ for symmetric unitary \gls{ris}.
		It is observed that both singular values can be manipulated up to $9\%$ using diagonal \gls{ris} and $42\%$ using symmetric \gls{bd}-\gls{ris}, despite both architectures have the same number of scattering elements and design parameters.
		A larger performance gap is expected when asymmetric \gls{bd}-\gls{ris} is available.
		This example shows \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values and motivates further studies on channel shaping.
	\end{subsection}

@@

\begin{section}{Simulation Results}
	In this section, we provide numerical results to evaluate the proposed \gls{bd}-\gls{ris} designs.
	Consider a distance-dependent path loss model $\Lambda(d) = \Lambda_0 d^{-\gamma}$ where $\Lambda_0$ is the reference path loss at distance \qty{1}{m}, $d$ is the propagation distance, and $\gamma$ is the path loss exponent.
	The small-scale fading model is $\mathbf{H} = \sqrt{\kappa/(1+\kappa)} \mathbf{H}_\text{LoS} + \sqrt{1/(1+\kappa)} \mathbf{H}_\text{NLoS}$, where $\kappa$ is the Rician $K$-factor, $\mathbf{H}_\text{LoS}$ is the deterministic \gls{los} component, and $\mathbf{H}_\text{NLoS} \sim \mathcal{CN}(\mathbf{0}, \mathbf{I})$ is the Rayleigh component.
	We set $\Lambda_0=\qty{-30}{dB}$, $d_\mathrm{D}=\qty{14.7}{m}$, $d_\mathrm{F}=\qty{10}{m}$, $d_\mathrm{B}=\qty{6.3}{m}$, $\gamma_\mathrm{D}=3$, $\gamma_\mathrm{F}=2.4$, $\gamma_\mathrm{B}=2$, $\kappa_\mathrm{D}=\kappa_\mathrm{F}=\kappa_\mathrm{B}=5$ for reference, which corresponds to an indoor environment with $\Lambda_\mathrm{D}=\qty{-65}{dB}$, $\Lambda_\mathrm{F}=\qty{-54}{dB}$, $\Lambda_\mathrm{B}=\qty{-46}{dB}$.

	\begin{subsection}{Channel Singular Values Redistribution}
		% \begin{figure}[!t]
		% 	\centering
		% 	\resizebox{0.65\columnwidth}{!}{
		% 		\input{assets/simulation/pc_singular_pareto.tex}
		% 	}
		% 	\caption{Singular value bounds for rank-1 indirect channel. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (4, 32, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (0, -17.5, -17.5) \unit{dB}$.}
		% 	\label{fg:pc_singular_pareto}
		% \end{figure}

		\begin{figure}[!t]
			\centering
			\subfloat[$2 \times 32 \times 2$ (no direct link)\label{fg:pc_singular_pareto_32_nd}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_pareto_32_nd.tex}
				}
			}
			\subfloat[$2 \times 32 \times 2$\label{fg:pc_singular_pareto_32}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_pareto_32.tex}
				}
			}
			\\
			\subfloat[$2 \times 64 \times 2$\label{fg:pc_singular_pareto_64}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_pareto_64.tex}
				}
			}
		\end{figure}

		\begin{figure}[!t]
			\centering
			\subfloat[$4 \times 32 \times 4$\label{fg:pc_singular_bound_32}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_bound_32.tex}
				}
			}
			\subfloat[$4 \times 64 \times 4$\label{fg:pc_singular_bound_64}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_bound_64.tex}
				}
			}
			\caption{
				Channel singular values with \gls{los} forward channel: analytical bounds (green lines) and optimized results (blue and red bars).
				`D' means diagonal \gls{ris} and `BD' means fully-connected \gls{bd}-\gls{ris}.
				% $\kappa_\mathrm{D}=\kappa_\mathrm{B}=0$, $\kappa_\mathrm{F}=\infty$.
			}
			% \caption{Singular values of equivalent channel with \gls{los} forward channel: analytical bounds (green lines) and optimized results (blue and red bars).}
			\label{fg:pc_singular_bound}
		\end{figure}

>>>>

\begin{section}{Channel Singular Values Redistribution}
	\begin{subsection}{A Toy Example}
		We first illustrate the channel shaping capabilities of different \gls{ris} by a toy example.
		% We first use a toy example to illustrate that \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
		Consider a $2 \times 2 \times 2$ setup where the direct link is blocked.
		The diagonal \gls{ris} is modeled by $\mathbf{\Theta}_\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{bd}-\gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}_\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		In particular, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		We also enforce symmetry by $\beta = \pi / 2$ such that both architectures have the same number of angular parameters.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/simulation/pc_singular_toy.eps}
			\caption{$2 \times 2 \times 2$ (no direct) channel singular value shaping by diagonal and symmetry unitary \gls{ris}.}
			\label{sm:pc_singular_toy}
		\end{figure}
		Fig.~\ref{sm:pc_singular_toy} shows the channel singular values achieved by an exhaustive grid search over $(\theta_1, \theta_2)$ for diagonal \gls{ris} and $(\alpha, \psi)$ for symmetric unitary \gls{ris}.
		It is observed that both singular values can be manipulated up to $9\%$ using diagonal \gls{ris} and $42\%$ using symmetric \gls{bd}-\gls{ris}, despite both architectures have the same number of scattering elements and design parameters.
		A larger performance gap is expected when asymmetric \gls{bd}-\gls{ris} is available.
		This example shows \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values and motivates further studies on channel shaping.
	\end{subsection}

@@

\begin{section}{Simulation Results}
	In this section, we provide numerical results to evaluate the proposed \gls{bd}-\gls{ris} designs.
	Consider a distance-dependent path loss model $\Lambda(d) = \Lambda_0 d^{-\gamma}$ where $\Lambda_0$ is the reference path loss at distance \qty{1}{m}, $d$ is the propagation distance, and $\gamma$ is the path loss exponent.
	The small-scale fading model is $\mathbf{H} = \sqrt{\kappa/(1+\kappa)} \mathbf{H}_\text{LoS} + \sqrt{1/(1+\kappa)} \mathbf{H}_\text{NLoS}$, where $\kappa$ is the Rician $K$-factor, $\mathbf{H}_\text{LoS}$ is the deterministic \gls{los} component, and $\mathbf{H}_\text{NLoS} \sim \mathcal{CN}(\mathbf{0}, \mathbf{I})$ is the Rayleigh component.
	We set $\Lambda_0=\qty{-30}{dB}$, $d_\mathrm{D}=\qty{14.7}{m}$, $d_\mathrm{F}=\qty{10}{m}$, $d_\mathrm{B}=\qty{6.3}{m}$, $\gamma_\mathrm{D}=3$, $\gamma_\mathrm{F}=2.4$ and $\gamma_\mathrm{B}=2$ for reference, which corresponds to a typical indoor environment with $\Lambda_\mathrm{D}=\qty{-65}{dB}$, $\Lambda_\mathrm{F}=\qty{-54}{dB}$, $\Lambda_\mathrm{B}=\qty{-46}{dB}$.
	Rayleigh fading is assumed for all channels $\kappa_\mathrm{D},\kappa_\mathrm{B},\kappa_\mathrm{F} \to \infty$ unless otherwise specified.

	\begin{subsection}{Channel Singular Values Redistribution}
		\begin{figure}[!t]
			\centering
			\subfloat[$2 \times 32 \times 2$ (no direct link)\label{fg:pc_singular_pareto_32_nd}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_pareto_32_nd.tex}
				}
			}
			\subfloat[$2 \times 32 \times 2$\label{fg:pc_singular_pareto_32}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_pareto_32.tex}
				}
			}
			\\
			\subfloat[$2 \times 64 \times 2$\label{fg:pc_singular_pareto_64}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_pareto_64.tex}
				}
			}
			\subfloat[$2 \times 128 \times 2$\label{fg:pc_singular_pareto_128}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_pareto_128.tex}
				}
			}
			\caption{Pareto frontiers of channel singular values reshaped by \gls{bd}-\gls{ris}.}
		\end{figure}

		\begin{figure}[!t]
			\centering
			\subfloat[$4 \times 32 \times 4$\label{fg:pc_singular_bound_32}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_bound_32.tex}
				}
			}
			\subfloat[$4 \times 64 \times 4$\label{fg:pc_singular_bound_64}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_bound_64.tex}
				}
			}
			\caption{
				Channel singular values with \gls{los} forward channel ($\kappa_\mathrm{F} = 0$): analytical bounds (green lines) and optimized results (blue and red bars).
				`D' means diagonal \gls{ris} and `BD' means fully-connected \gls{bd}-\gls{ris}.
				% $\kappa_\mathrm{D}=\kappa_\mathrm{B}=0$, $\kappa_\mathrm{F}=\infty$.
			}
			% \caption{Singular values of equivalent channel with \gls{los} forward channel: analytical bounds (green lines) and optimized results (blue and red bars).}
			\label{fg:pc_singular_bound}
		\end{figure}
